{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-10-27 thesis/education.tex
final pre-defense draft revisions

The data set is comprised of Landslide's automatically-generated usage snapshots
from CMU students from semesters Spring 2015 to Spring 2018 inclusive,
% and PSU \psuos from Spring 2018, % no snapshots from psu
CMU's official project grades from same,
CMU students' survey responses and submitted project code from Fall 2017 and Spring 2018
PSU students' survey responses and submitted project code from Spring 2018,
and U. Chicago students' survey responses from Fall 2017.
Table~\ref{tab:photo-of-ze-studence} shows the student participation rate across semesters
% TODO update for psu
for the thread library study%
\footnote{
Participation at CMU was determined by receiving any automatic usage snapshots;
participation at PSU was determined, for lack of snapshots, by sign-up form submissions, which might over-count slightly.
}.
% and PSU

\begin{table}[h]
	\begin{center}
		% ugh this looks so stupid
		% but better than latex's output looking stupid
		\begin{tabular}{r|crcrc|r}
			& \multicolumn{5}{c|}{\bf Used Landslide?} & \\
			\bf Semester & & \bf yes & & \bf no & & \bf Total \\
			\hline
			CMU S'15	& &   8	& &  18	& &  26 \\
			CMU F'15	& &  22	& &   8	& &  30 \\
			CMU S'16	& &  18	& &  16	& &  34 \\
			CMU F'16	& &  12	& &   6	& &  18 \\
			CMU S'17	& &  19	& &  13	& &  32 \\
			CMU F'17	& &  14	& &   5	& &  19 \\
			CMU S'18	& &  10	& &   8	& &  18 \\
			\hline
			U. Chicago F'17	& &   4	& &  17	& &  21 \\ % 21 taken from number of submitted 'threads', idk abt userprog
			% TODO when you adjust this, adjust the % of participance down in the discussion para
			PSU S'18	& &  38	& &  98	& & 136 \\
			\hline
			\bf CMU total	& & 103	& &  74	& & 177 \\
			\bf Non-CMU total
					& &  42	& & 115	& & 157 \\
			\bf Total	& & 145	& & 189	& & 327 \\
		\end{tabular}
	\end{center}
	\caption{Participation rates across semesters,
	among students who submitted thread libraries (CMU, PSU) or kernels (U. Chicago).}
	\label{tab:photo-of-ze-studence}
\end{table}

>>>>

The data set is comprised of Landslide's automatically-generated usage snapshots
from CMU students from semesters Spring 2015 to Spring 2018 inclusive,
% and PSU \psuos from Spring 2018, % no snapshots from psu
CMU's official project grades from same,
CMU students' survey responses and submitted project code from Fall 2017 and Spring 2018
PSU students' survey responses and submitted project code from Spring 2018,
and U. Chicago students' survey responses from Fall 2017.
Table~\ref{tab:photo-of-ze-studence} shows the student participation rate across semesters
for the thread library study%
\footnote{
Participation at CMU was determined by receiving any automatic usage snapshots;
participation at PSU was determined, for lack of snapshots, by sign-up form submissions, which might over-count slightly.
}.
% and PSU

\begin{table}[h]
	\begin{center}
		% ugh this looks so stupid
		% but better than latex's output looking stupid
		\begin{tabular}{r|crcrc|r}
			& \multicolumn{5}{c|}{\bf Used Landslide?} & \\
			\bf Semester & & \bf yes & & \bf no & & \bf Total \\
			\hline
			% group-[redacted] had a p2/ dir, but not in grade files
			CMU S'15	& &   8	& &  18	& &  26 \\
			CMU F'15	& &  22	& &   8	& &  30 \\
			CMU S'16	& &  18	& &  16	& &  34 \\
			CMU F'16	& &  12	& &   6	& &  18 \\
			CMU S'17	& &  19	& &  13	& &  32 \\
			CMU F'17	& &  14	& &   5	& &  19 \\
			CMU S'18	& &  10	& &   8	& &  18 \\
			\hline
			U. Chicago F'17	& &   4	& &  17	& &  21 \\ % 21 taken from number of submitted 'threads', idk abt userprog
			PSU S'18	& &  38	& &  98	& & 136 \\
			\hline
			\bf CMU total	& & 103	& &  74	& & 177 \\
			\bf Non-CMU total
					& &  42	& & 115	& & 157 \\
			\bf Total	& & 145	& & 189	& & 327 \\
		\end{tabular}
	\end{center}
	\caption{Participation rates across semesters,
	among students who submitted thread libraries (CMU, PSU) or kernels (U. Chicago).}
	\label{tab:photo-of-ze-studence}
\end{table}

### 2018-10-27 thesis/intro.tex
final pre-defense draft revisions

\begin{center}
	{\em Combining both theoretically-founded automatic reduction techniques
	and user-informed heuristic ones,
	stateless model checking can sufficiently mitigate exponential explosion
	to be a practical testing technique even for inexperienced users.}
\end{center}

>>>>

\begin{center}
	{\em Combining %both
	theoretically-founded automatic reduction techniques
	and user-informed heuristic ones,
	stateless model checking can sufficiently mitigate exponential explosion
	to be a practical testing technique for inexperienced users and real-world programs alike.}
\end{center}

### 2018-10-27 thesis/landslide.tex
final pre-defense draft revisions

Landslide detects such ad-hoc {\em yield-loop blocking}
by keeping a counter for each thread to track how many {\tt yield}s it has invoked since the last ``interesting'' activity.
Interesting here is defined heuristically as any other known P2 API call,
with the exception of {\tt mutex\_lock()} and {\tt mutex\_unlock()}%
\footnote{Landslide must recognize yield-blocking loops that contain mutex operations
to allow for open-coded reimplementations of {\tt cond\_wait()},
% see comment above "USER_MUTEX_YIELD_ACTIVITY" in user sync.c
which for example the {\tt paraguay} test uses intentionally,
because it is testing the correctness of the student's {\tt cond\_wait()}.}.
Whenever this counter reaches the heuristic cutoff of 10 ({\tt TOO\_MANY\_YIELDS}),
Landslide declares the thread blocked,
and treats it just as if it had invoked {\tt deschedule()} for purposes of DPOR
({\tt check\_user\_yield\_activity()} in {\tt user\_sync.c}).
% this also involves propagating yield-blockedness back up the branch, along those 10 pps where it was still counting up to 10
% and also refusing to run any other threads while yield is going on - KEEP_RUNNING_YIELDING_THREADS
%
In cases where {\tt yield} itself is not involved,
Landslide also counts the number of atomic instructions ({\tt xchg}, {\tt xadd}, {\tt cmpxchg}, etc),
and marks the thread blocked if it exceeds 100 such ({\tt TOO\_MANY\_XCHGS\_TIGHT\_LOOP})%
\footnote{If preemption points exist in between, instead only 20 such ({\tt TOO\_MANY\_XCHGS\_WITH\_PPS}),
to avoid stressing DPOR's $O(n)$ independence computation.}.
When such a loop contains neither {\tt yield} nor atomics,
Landslide falls back on the standard infinite loop detector and reports a bug directly, as described above.
Future work could extend this to include more modern ways of establishing memory safety such as
acquire/release barriers
%\cite{sully-thesis}
and hardware transactions.

@@

Landslide will retry this process up to a heuristic limit of 128 times ({\tt DEADLOCK\_FP\_MAX\_\allowbreak()ATTEMPTS}
before issuing a deadlock bug report.
The overhead of this check is linear time in the number of retries permitted,
although this is typically negligible compared to the exponential size of state spaces overall.
If the heuristic limit is too small, very ``loopy'' programs
(which invoke xchg or yield therein)
could falsely exhaust this limit while not being truly deadlocked,
so a good limit should be well higher than the total number of concurrency events expected for any Landslide-friendly test.
The cost of a high limit manifests in the length of preemption traces when deadlock is declared,
which will display a proportional number of meaningless preemption points,
although future work could easily truncate them retrospectively.
This check is implemented in {\tt try\_avoid\_fp\_deadlock()} in {\tt arbiter.c}.

>>>>

Landslide detects such ad-hoc {\em yield-loop blocking}
by keeping a counter for each thread to track how many {\tt yield}s it has invoked since the last ``interesting'' activity.
% nb. not system calls, not even d/mr, as some student mutexes may use those and you don't want to count them
Interesting here is defined heuristically as any other known P2 API call,
with the exception of {\tt mutex\_lock()} and {\tt mutex\_unlock()}%
\footnote{Landslide must recognize yield-blocking loops that contain mutex operations
to allow for open-coded reimplementations of {\tt cond\_wait()},
% see comment above "USER_MUTEX_YIELD_ACTIVITY" in user sync.c
which for example the {\tt paraguay} test uses intentionally,
because it is testing the correctness of the student's {\tt cond\_wait()}.}.
Whenever this counter reaches the heuristic cutoff of 10 ({\tt TOO\_MANY\_YIELDS}),
Landslide declares the thread blocked,
and treats it just as if it had invoked {\tt deschedule()} for purposes of DPOR
({\tt check\_user\_yield\_activity()} in {\tt user\_sync.c}).
% this also involves propagating yield-blockedness back up the branch, along those 10 pps where it was still counting up to 10
% and also refusing to run any other threads while yield is going on - KEEP_RUNNING_YIELDING_THREADS
%
In cases where {\tt yield} itself is not involved,
Landslide also counts the number of atomic instructions ({\tt xchg}, {\tt xadd}, {\tt cmpxchg}, etc),
and marks the thread blocked if it exceeds 100 such ({\tt TOO\_MANY\_XCHGS\_TIGHT\_LOOP})%
\footnote{If preemption points exist in between, instead only 20 such ({\tt TOO\_MANY\_XCHGS\_WITH\_PPS}),
to avoid stressing DPOR's $O(n)$ independence computation.}.
When such a loop contains neither {\tt yield} nor atomics,
Landslide falls back on the standard infinite loop detector and reports a bug directly, as described above.
Future work could extend this to include more modern ways of establishing memory safety such as
acquire/release barriers
%\cite{sully-thesis}
and hardware transactions.

@@

Landslide will retry this process up to a heuristic limit of 128 times ({\tt DEADLOCK\_FP\_MAX\_\allowbreak{}ATTEMPTS}
before issuing a deadlock bug report.
The overhead of this check is quadratic time in the number of retries permitted
(as DPOR is quadratic in the overall branch depth),
although this is negligible compared to the exponential size of state spaces overall.
If the heuristic limit is too small, very ``loopy'' programs
(which invoke xchg or yield therein)
could falsely exhaust this limit while not being truly deadlocked,
so a good limit should be well higher than the total number of concurrency events expected for any Landslide-friendly test.
The cost of a high limit manifests in the length of preemption traces when deadlock is declared,
which will display a proportional number of meaningless preemption points,
although future work could easily truncate them retrospectively.
This check is implemented in {\tt try\_avoid\_fp\_deadlock()} in {\tt arbiter.c}.
